\toolname: Property-Based Testing for MPC Security

### 2023-11-03 paper/main.tex

\subsection{Formal-methods for cryptography}
Although a useful language-driven test for MPC security remains out of reach,
a variety of useful work aproaches the subject from different directions.

Haagh et al.~\cite{haagh2018computer} % Computer-aided proofs for multiparty computation with active security
use EasyCrypt to manually show the security of GMW.
Their definitions generalize to protocols with efficiently invertible functionalities.
They model MPC as a non-interference property: Honest secrets are non-interfering on the corrupt views conditioned on the corrupt outputs.
If their system could be automated and paired with language constraints to ensure invertibility,
then it could apply to many, but certainly not all, use cases.

Barthe et al.~\cite{barthe2019probabilistic} % https://arxiv.org/abs/1907.10708 A Probabilistic Separation Logic
give a framework for probabilistic reasoning and show that it can be used to prove security of additive secret sharing.
Even if the technique could be automated, it's unclear how it would generalize to any other kind of protocol.

$\lambda_\textbf{obliv}$~\cite{darais2019language} %
gives a type system that ensures uniform and independent randomness of key values throughout a program.
Their system could be adapted to ensure perfect uniformity of parties' views in a protocol,
but this would of course prevent any non-vacuous functionality from being implemented.

\textsc{Owl}~\cite{gancher2023owl} % https://eprint.iacr.org/2023/473.pdf
gives a type system guaranteeing non-interference with an (extensible) selection of cryptographic operations as language primitives.
They can type-check convoluted uses of operations like encryption and hashing, but without features for inference or inversion,
there's no way to verify an implementation of MPC.

Fournet et al.~\cite{fournet2011information} % https://www.microsoft.com/en-us/research/wp-content/uploads/2017/01/information-flow-types-for-homomorphic-encryptions-ccs11.pdf
also use a type system to ensure safe usage of cryptographic primitives,
but they specifically work with homomorphic encryption schemes and the various measures by which such schemes may be correct.
Like OWL, their type system assumes a secure implementation of the encryption scheme.

We experimented with choreography languages that translate directly into Sum Product Networks (SPNs)~\cite{poon2011sum}.
A SPN is a directed-graph representation of a multi-variate distribution.
They are compact, allow efficient inference, and can represent values symbolically as needed,
so testing for conditional independence is practical.
However, such a language cannot express any existing MPC protocols, and likely cannot express any \textit{useful} protocols,
because if the language allowed any value to be used more than once, it would result in graphs that violate key invariants of SPNs.

The \textsc{Dice} language~\cite{holtzen2020scaling} translates programs into binary decision diagrams (BDDs),
analogous to the way we tried to translate them into SPNs.
BDDs can represent, and do inference on, distributions that aren't representable as SPNs,
but since the graph \textit{structure} of the BDD of a \textsc{Dice} program depends on the actual \textit{values} on which one is conditioning,
a \textsc{Dice}-based test for the kind of conditional independence needed for MPC would scale exponentially with the input/output width of the protocol.

Lilac~\cite{li2023lilac} % https://arxiv.org/abs/2304.01339
is a program logic for reasoning about probability distributions and PPLs.
It more expressive than the original probabilistic separation logic~\cite{barthe2019probabilistic},
but it has not yet been applied to MPC, nor is it clear how it would be used in an automated test.

@@

Our \toolname python script consumes a stream of fixed-length bit vectors as a CSV.
Annotation of which columns are available in the real and ideal worlds is provided in the header.
The CSV is parsed and broken up as needed as a \texttt{pandas} dataframe.
The individulal ML models are instances of \texttt{sklearn.tree.DecisionTreeClassifier},
and they're trained in parallel with \texttt{joblib.Parallel}.
The D-Tree scores are inflated by $+1 \cdot 10^{-10}$ to avoid division-by-zero in the score-ratio,
and the score distributions are compared by \texttt{scipy.stats.wilcoxon}.

@@

>>>>

@@

Our \toolname Python script consumes a stream of fixed-length bit vectors as a CSV.
Annotation of which columns are available in the real and ideal worlds is provided in the header.
The CSV is parsed and broken up as needed as a \texttt{pandas} dataframe.
The individual ML models are instances of \texttt{sklearn.tree.DecisionTreeClassifier},
and they're trained in parallel with \texttt{joblib.Parallel}.
The D-Tree scores are inflated by $+1 \cdot 10^{-10}$ to avoid division-by-zero in the score-ratio,
and the score distributions are compared by \texttt{scipy.stats.wilcoxon}.

@@

\subsection{Formal-methods for cryptography}
Although automated language-based verification of MPC security is an open question,
a variety of useful work approaches the subject from different directions.

Haagh et al.~\cite{haagh2018computer} % Computer-aided proofs for multiparty computation with active security
use EasyCrypt to manually show the security of GMW.
Their definitions generalize to protocols with efficiently invertible functionalities.
They model MPC as a non-interference property: Honest secrets are non-interfering on the corrupt views conditioned on the corrupt outputs.
Their approach requires manual proof in EasyCrypt.

Barthe et al.~\cite{barthe2019probabilistic} % https://arxiv.org/abs/1907.10708 A Probabilistic Separation Logic
give a framework for probabilistic reasoning and show that it can be used to prove security of additive secret sharing.
However, the framework is not amenable to automation.

$\lambda_\textbf{obliv}$~\cite{darais2019language} %
gives a type system that ensures uniform and independent randomness of key values throughout a program.
Their system is automated, and could be adapted to ensure perfect noninterference between honest secrets and corrupt views,
but does not support the kind of \emph{conditional} independence required for most MPC protocols.

\textsc{Owl}~\cite{gancher2023owl} % https://eprint.iacr.org/2023/473.pdf
gives a type system guaranteeing non-interference with an (extensible) selection of cryptographic operations as language primitives.
OWL is designed for secure protocols that reveal nothing to the adversary, and does not support verification of MPC security.

Fournet et al.~\cite{fournet2011information} % https://www.microsoft.com/en-us/research/wp-content/uploads/2017/01/information-flow-types-for-homomorphic-encryptions-ccs11.pdf
also use a type system to ensure safe usage of cryptographic primitives,
but they specifically work with homomorphic encryption schemes and the various measures by which such schemes may be correct.
Like \textsc{Owl}, their type system assumes a secure implementation of the encryption scheme and does not support verification of MPC security.

Several PPLs directly support efficient inference and some forms of independence testing.
A Sum Product Network (SPN)~\cite{poon2011sum} is a directed-graph representation of a multi-variate distribution;
SPPL~\cite{saad2021sppl} is a PPL that performs inference using SPNs.
SPNs are compact, allow efficient inference, and can represent values symbolically as needed,
so testing for conditional independence is practical.
However, such a language cannot express any existing MPC protocols, and likely cannot express any \textit{useful} protocols,
because if the language allowed any value to be used more than once, it would result in graphs that violate key invariants of SPNs.

The \textsc{Dice} language~\cite{holtzen2020scaling} analogously translates programs into binary decision diagrams (BDDs).
BDDs can represent, and do inference on, distributions that aren't representable as SPNs,
but since the graph \textit{structure} of the BDD of a \textsc{Dice} program depends on the actual \textit{values} on which one is conditioning,
a \textsc{Dice}-based test for the kind of conditional independence needed for MPC would scale exponentially with the input/output width of the protocol.

\textsc{Lilac}~\cite{li2023lilac} % https://arxiv.org/abs/2304.01339
is a program logic for reasoning about probability distributions and PPLs.
It more expressive than the original probabilistic separation logic,
but does not support automation.
